---
title: "Data 607 - Week 3 Assignment"
author: "VINAYAK PATEL"
date: "September 16, 2018"
output: html_document
css: style.css
---

#### **Question 3. Copy the introductory example. The vector name stores the extracted names.**

<font size="5" style="color:#1462e0;"> a. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.</font> 
 
```{r}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
#extract raw data to get names
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names

#extract data to get phone number
phone <- unlist(str_extract_all(raw.data,"\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone

#To remove Title and Middle initial 
temp_name<- unlist(str_replace_all(names, "[[:alpha:]]+\\. ", ""))

#To Remove Punctuation in name
Full_name<-str_replace_all(temp_name, "[[:punct:]]", "")

Full_name
```

**I am tring to separate First_name and Last_name into different columns  **

```{r}
#Split Full name
temp <- strsplit(Full_name, " ")
temp
#Create Matrix 
mat  <- matrix(unlist(temp), ncol=2, byrow=TRUE)

mat
#convert Matrix into Data.Frame
df   <- as.data.frame(mat)

df
#combine with phone number

df   <- cbind(df, phone)
#give valid name to column name
colnames(df) <- c("First_name", "Last_name", "Phone")
df
```


<font size="5" style="color:#1462e0;"> b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.) </font>

    
```{r}
title <- str_detect(names, "Rev.|Dr.")
title
```

<font size="5" style="color:#1462e0;"> c. Construct a logical vector indicating whether a character has a second name.</font>
    
```{r}
initial <- str_replace(names, "[[:alpha:]]{2,3}[.]", "")
initial <- str_count(initial, "\\w+")
initial <- initial > 2;
initial

```

<hr>

#### **Question 4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.**

<font size="5">
**1. [0-9]+\\$ **

Numeric characters only (0 to 9), any length >0, has to end with a "$" sign. 

```{r}
str_detect("888888$", "[0-9]+\\$")
```

<font size="5">
**2. \\b[a-z]{1,4}\\b** 

Vectors containing strings with maximum 4 continuous characters lowercase letters from a to z at the beginning and at the end.

```{r, include=TRUE}
str_detect("abcdefgh", "\\b[a-z]{1,4}\\b")

```

<font size="5">
3. .*?\\.txt$

this will return any string with any length that ends with '.txt'  
   
```{r, include=TRUE}
str_detect("Sample1.txt", ".*?\\.txt$")

```
<font size="5">
**4.\\d{2}/\\d{2}/\\d{4}**

Has to contain a string with: 2 digits, followed by string "/", followed by 2 digits, followed by string "/", and then followed by 4 digits. most likely date format e.g dd/mm/yyyy or mm/dd/yyyy.
    
```{r, include=TRUE}
str_detect("16/09/2018", "\\d{2}/\\d{2}/\\d{4}")
```
<font size="5">
**5. <(.+?)>.+?</\\1> **

This will helps to backreferencing to return any string with in HTML Tags. 

```{r, include=TRUE}
tag1 <- "<<h3>This is a heading</h3>
  <p>This is a paragraph.</p>"

str_extract_all(tag1, '<(.+?)>.+?</\\1>')
```

9. Secret Message
The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

**clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo**
**Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO**
**d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5**
**fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr**
```{r}
secret_message<- paste("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
#get all uppercase letters.
message <- unlist(str_extract_all(secret_message, "[[:upper:].]{1,}"))
#join all letters.
message <- str_replace_all(paste(message, collapse = ''), "[.]", " "); 
message

```

