---
title: 'Data 607: Project 3'
author: "VINAYAK PATEL"
date: "October 19, 2018"
output:
  html_document:
    collapsed: FALSE
    theme: cosmo
    toc: yes
    toc_float: no
    toc_depth: 2
    df_print: paged

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Environment Prep
```{r, warning=FALSE, message=FALSE}
if (!require('rvest')) install.packages('rvest')
if (!require('XML')) install.packages('XML')
if (!require('jsonlite')) install.packages('jsonlite')
if (!require('RCurl')) install.packages('RCurl')
if (!require('rlist')) install.packages('rlist')
if (!require('magrittr')) install.packages('magrittr')
if (!require('tidyverse')) install.packages('tidyverse')
if (!require('RJSONIO')) install.packages('RJSONIO')
if (!require('DT')) install.packages('DT')

```

# Abstract

##Context
 
<br>   For the project 3, we have decided to conduct survey to find out the necessary soft-skills, and technical skills. Which programming language is need to learn first, and which programming language is necessary to be successful Data Scientist.

<br>   Kaggle already has the wide survey available for the Data science and Machine learning with more than 16000 respondents but data scientist also needs to know how to conduct survey in order to generate own datasets. 

# Data Source

##Methodology
<br>  This survey received 72 respondents from around the world. We will some excluded respondents who answered "No" to the question regarding their IT professional because these respondents can affect the overall analysis. even though this question was the first required question, so not answering it indicates that the respondent did not proceed further in survey and redirected to end of the survey.

<br> The survey was taken during October 17 to October 20th. Most of the questions was not required because they can submit most information survey as many as they can and didn't forced to select unnecessary and wrong options. 

## platform used
This survey was conducted using Google Forms. Also, most of the respondents were from LinkedIn group: "Data Scientist & Analyst", "Database Administrator (+5000 DBAs from around the world) Join now!", and "Datacamp Community" Slack Channel. 


# Research Question

<br>

## **Which are the most valued data science skills?**  

# Importing data
```{r}

# Import survey data


rawData <- read.csv("https://raw.githubusercontent.com/Vinayak234/SPS_DATA_607/master/SPS_DATA_607/Project_3/To%20find%20Most%20Valued%20Data%20Science%20Skills.csv", stringsAsFactors = TRUE, header = TRUE)

rawData

# Dropping none IT professionals records
clean_rawData<- rawData[!grepl("No", rawData$Are.you.related.to.IT.Profession.),]

#Drop column not needed
drops_c <- c("Timestamp","Are.you.related.to.IT.Profession.")
clean_rawData<-clean_rawData[ , !(names(clean_rawData) %in% drops_c)]

head(clean_rawData)

#Short name for column name
Short_name <- c("Gender","Country","Age","Are you in School","Eduction","Major","Are you learning data science skill","Emplyment Status","Title","Experience","Blogs","College","Projects", "Online Course", "Friends","Co-workers","Youtube","Textbooks","First language","must language","Amazon Machine Learning","Big Data","College Degree","Data Visualizations","Enterprise Tools","Google Cloud","Hadoop/Hive/Pig","IBM SPSS","Java","Microsoft Excel","NoSQL","Oracle Data Mining","Python","R","Relational data","SAS","SQL","Tableau",
"Intellectual curiosity","Business acumen", "Communication skills","Teamwork","Collaboration","Creative Thinking","Problem Solving", "Active Learning","Perceptiveness","Interpersonal Skills","Generating Hypotheses")

colnames(clean_rawData) <- Short_name
head(clean_rawData)


#VIEW THE DATA AND CREATE NAMES
summary(clean_rawData)



```

## Gender, Demograph, and Age Information

```{r message=FALSE}

#to see how many of the participants were male or female by their age
DataGender <- clean_rawData%>%
  select(Gender, Age) %>%
  filter(trimws(clean_rawData$Country)!='',Age!='NA') %>%
  group_by(Age,Gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

DataGender

```


```{r}
# Plot
library(ggplot2)
ggplot(data = DataGender, 
       mapping = aes(x = Age, fill = Gender, 
                     y = ifelse(test = Gender == "Female", 
                                yes = -count, no = count)))+
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = abs) +
  labs(y="Count") 

```

```{r}
DemoGraphicsData <- clean_rawData %>%
  group_by(Country)%>% # Group by country
  summarise(count=n()) %>%# Count how many respondents selected each option
  mutate(percent = (count / sum(count)) * 100) %>%     
  arrange(desc(count))# Arrange the counts in descending order
 
DemoGraphicsData

#plot graph
DemoGraphicsData[] %>%
  arrange(count) %>%
  mutate(Country=factor(Country, levels=Country)) %>%
  ggplot( aes(x=Country, y=count)) +
    geom_segment( aes(xend=Country, yend=0)) +
    geom_point( size=2, color="orange") +
    coord_flip() + labs(title="plot by Country")+ 
    theme_minimal()

```




```{r}
# to findout most recommanded language

clean_rawData_First_Language <- clean_rawData %>%  
    # Remove any rows where the respondent didn't answer the question
    filter(trimws(clean_rawData$`First language`)!='')%>%
    # Group by the responses to the question
    group_by(`First language`) %>% 
    # Count how many respondents selected each option
    summarise(count = n()) %>% 
    # Calculate what percent of respondents selected each option
    mutate(percent = (count / sum(count)) * 100) %>% 
    # Arrange the counts in descending order
    arrange(desc(count))

clean_rawData_First_Language

# to findout have to learn language

clean_rawData_Neccecery_Language <- clean_rawData %>%  
    # Remove any rows where the respondent didn't answer the question
  filter(trimws(clean_rawData$`must language`)!='')%>%
  # Group by the responses to the question
  group_by(`must language`) %>% 
  # Count how many respondents selected each option
  summarise(count = n()) %>% 
  # Calculate what percent of respondents selected each option
  mutate(percent = (count / sum(count)) * 100) %>% 
  # Arrange the counts in descending order
  arrange(desc(count))

clean_rawData_Neccecery_Language

```

## To find out helpfulness of Learning way 
```{r}
clean_rawData_helpful <- clean_rawData %>%
  select(Blogs:Textbooks)

clean_rawData_helpful


clean_rawData_helpful$id <- seq.int(nrow(clean_rawData_helpful))

clean_rawData_helpful <- gather(clean_rawData_helpful, platforms, How.useful,
                                Blogs:Textbooks) %>%
  arrange(id) %>%
  group_by(platforms) %>%
  filter(How.useful!='NA', How.useful!='')

clean_rawData_helpful

```

```{r fig.width=10, fig.height=5}

ggplot(data = clean_rawData_helpful,aes(fill=How.useful, platforms, How.useful))+
  geom_bar(stat = 'identity')+  
    labs(title = "How useful these platforms are for New data scientist", 
                         x = "Platforms ", 
                         y = "How useful") +
theme_bw()+
theme(axis.text.x = element_text(),axis.text.y=element_blank(),axis.ticks.y=element_blank())
```


## Result 
Analysis on


## Third





